---
title: "A3: Incarceration Report"
author: "Mike Ramsay"
date: "2/25/2022"
output: html_document
---

```{r setup, include=FALSE}
source("../source/analysis.R")

library("knitr")
```


## Introduction
For this assignment, I am looking at incarceration statistics in the US during the past couple decades. I used a dataset from the Vera institute that contains lots of statistics on total populations, by state and county and their breakdowns by race and sex. It also contains information about the people incarcerated, like their race, whether they are a juvenile and why they are in jail. Some of the specific values that I explored are the states that have the largest proportions of their populations in jails as well as the total jail population changes over the past couple decades and what the jail population looks like when broken down by race.

## Summary Information

```{r summaryInfo, echo = FALSE}
# Summary Information
average_jail_pop <- av_jail_pop
largest_jail_pop <- state_largest_jail_pop
largest_juvenile <- state_largest_juvenile_pop
jail_pop_change_1980_to_2000 <- jail_pop_1980_to_2000
proportion_of_blacks_jail <- prop_blacks
```

The summary information that I collected, includes some values that I think are interesting or important to this dataset. I found that the average of all jail populations across the US during 2018 was **`r round(average_jail_pop)`**. The state with the largest jail population, _proportional_ to its population is **`r largest_jail_pop`** and the state with the largest juvenile jail population _proportional_ to its population is **`r largest_juvenile`**. Also the change in jail population across the United States, from 1980 to 2000 is an increase of **`r round(jail_pop_change_1980_to_2000)`** prisoners. The proportion of black people in jails around the United States, from the year 2000 to 2013, is **`r round(proportion_of_blacks_jail)`**%.


## Charts

### Trends Over Time Chart

```{r time_plot, echo=FALSE}
plot(time_plot)
```

For this chart, I decided on measuring the total jail population in California from 2000 to 2013. I decided to include this chart because I was interested in what the most populus state's jail population looked like. I also think that this chart shows some interesting trends, like how the total population saw a spike around 2008 with just over 50,000 people in jail.

### Variable Comparison Chart

```{r var_comparison_plot, echo=FALSE}
plot(var_comparison_plot)
```

This is a chart comparing the Latinx population and the total population of people who are in jail from ICE. This is using the total Latinx population aged 15 to 64 because the Vera institute mentioned that this is the age group that is most likely to be jailed. It also shows a positive relationship, between counties with greater Latinx populations and more people being imprisoned from ICE. I included this chart because I think it shows some valuable information about some of the current issues with ICE's role in our country and how some of these people of Latinx decent may be unfairly treated or targeted. One thing to add it that I only selected counties with a Latinx Population of less than one million. I did this to allow for more detail and make the graph look better, because these counties, most of which were New York, were way off in the upper right (as expected). 

### Map

```{r us_plot, echo=FALSE}
plot(us_plot)
```

This is a map of the United States. 

